---
title: "Data Collection and Formatting"
author: "Group 33: Greg Foral, Seung Woo Choi, Eduardo Arias Villanueva, Tyler Jeron Lang"
date: '2022-10-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(dplyr)
```

## Introduction

For ease of analysis, combining all the seperate datasets into one 'master' dataset is ideal. To begin, we start by importing data from [Zillow Housing Data](https://www.zillow.com/research/data/)

```{r imports}
#contains the number of homes for sale
inventory_for_sale_df <- read.csv('./data/inventory_for_sale.csv', header=TRUE)

#contains the median list prices of homes for a region
median_list_price_df <- read.csv('./data/median_list_price.csv', header=TRUE)

#contains median sale prices of homes for a region
median_sale_price_df <- read.csv('./data/median_sale_price.csv', header=TRUE)

#contains median sale price to list price ratio of homes for a region
median_sale_to_list_ratio_df <- read.csv('./data/median_sale_to_list_ratio.csv', header=TRUE)

#contains percent of homes sold above list price for a region
prct_homes_above_list_df <- read.csv('./data/prct_homes_above_list.csv', header=TRUE)

#contains the Zillow Home Value Index (ZHVI) for a region. This value is calculated by Zillow to estimate the value of a home within the 35th to 65th percentile range
zhvi_df <- read.csv('./data/ZHVI.csv', header=TRUE)
```

## Initial Inspection of Data

We can now preview the initial dataframes and inspect if any changes need to be made:

```{r inspect}
head(inventory_for_sale_df, 5)
```

A couple things to note here that apply to all of the datasets imported above - the **date** columns are now prefixed with 'X' and follow the format Xmm.dd.yyyy. This is different from the actual format within the .csv files - the original **date** columns contained the '/' character which would result in an invalid column name, so R automatically converts these columns into a valid format. We will clean up these values later on in this notebook.

Ideally, we can plan our 'master' dataset to contain the following schema:

* Region ID
* SizeRank
* RegionName
* RegionType
* StateName
* Latitude
* Longitude
* Month
* Inventory_for_sale
* Median_list_price
* Median_sale_price
* Median_StL_ratio
* Prct_homes_above_list
* ZHVI

To follow this schema, we first have to reshape this dataframes from wide to long format:

```{r reshape}
inventory_for_sale_df <- inventory_for_sale_df %>%
  gather(month, inventory, -c(RegionID, SizeRank, RegionName, RegionType, StateName))

median_list_price_df <- median_list_price_df %>%
  gather(month, median_list_price, -c(RegionID, SizeRank, RegionName, RegionType, StateName))

median_sale_price_df <- median_sale_price_df %>%
  gather(month, median_sale_price, -c(RegionID, SizeRank, RegionName, RegionType, StateName))

median_sale_to_list_ratio_df <- median_sale_to_list_ratio_df %>%
  gather(month, StL_ratio, -c(RegionID, SizeRank, RegionName, RegionType, StateName))

prct_homes_above_list_df <- prct_homes_above_list_df %>%
  gather(month, prct_above_list, -c(RegionID, SizeRank, RegionName, RegionType, StateName))

zhvi_df <- zhvi_df %>%
  gather(month, ZVHI, -c(RegionID, SizeRank, RegionName, RegionType, StateName))

```

Now we join the dataframes together:

```{r merge}
join_cols <- c('RegionID', 'SizeRank', 'RegionName', 'RegionType', 'StateName', 'month')

main_df <- full_join(zhvi_df, inventory_for_sale_df, by=join_cols) %>%
  full_join(., median_list_price_df, by=join_cols) %>%
  full_join(., median_sale_price_df, by=join_cols) %>%
  full_join(., median_sale_to_list_ratio_df, by=join_cols) %>%
  full_join(., prct_homes_above_list_df, by=join_cols)

head(main_df, 5)
```

It's important to consider the ZHVI dataset's date range, which goes back to Jan 2000 - much earlier than the other data sources. Therefore, there are many NULL values for the remaining variables.

Next, we can use the [US Cities](https://www.kaggle.com/datasets/sergejnuss/united-states-cities-database?resource=download) dataset to join in location (lat / long) information:

```{r geocode}

#bring in the uscities location data, and concatinate the city / state columns to match
#the formatting of the main_df
location_df <- read.csv('./data/uscities.csv') %>%
  select(city, state_id, lat, lng) %>%
  unite(RegionName, c('city', 'state_id'), sep = ', ')

main_df <- left_join(main_df, location_df, by='RegionName')
```

The last thing to do to prepare our dataset is to format the 'month' column, which is still prefixed with the 'X' character:

```{r dateformat}
main_df$month <- as.Date(main_df$month, format = "X%m.%d.%Y")

head(main_df, 5)
```

Now that the date column is formatted, we are ready to start analyzing the data.

```{r write}
write.csv(main_df, './data/formatted_df.csv')
```
