---
title: "preliminary_eda-and-data-cleaning"
output: html_document
date: "2022-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Preliminary EDA

Let's start by clearing the environment and setting a random seed to allow for code reproducibility.

```{r env}
# Clear the environment
rm(list = ls())

# Set a random seed for code reproducibility
set.seed(25)
```

Let's install any required libraries.

```{r libraries, results = 'hide'}
# Install required libraries
install.packages("xts", repos = "http://cran.us.r-project.org")
library(xts)

install.packages("lubridate", repos = "http://cran.us.r-project.org")
library(lubridate)

library(ggplot2)
```

Let's retrieve the formatted csv file: formatted_df.csv.

```{r data}
# Read the formatted csv file
data <- read.csv(file = "../0_Data Collection and Formatting/data/formatted_df.csv")

# View the structure of the data
str(data)
```

Looking at the data, I notice two main things: (1) some variables require a data type conversion (e.g. RegionID and month) and (2) some variables have lots of missing values.\

Let's handle the data type conversions first.

```{r dtype conversion}
# Convert RegionID from int to chr
data$RegionID <- as.character(data$RegionID)
typeof(data$RegionID)

# Create a new variable, ts_month, with data type "Date"
data$ts_month <- ymd(data$month)

# View the dataframe with ts_month
data <- data[order(data$ts_month), ]
str(data)
```

The data type conversions are complete for now.\

Let's explore the missing values.

```{r missing values}
# Create a list of column names
col_names <- colnames(data)

# Define a function to print the number of missing values for all variables
missing_values <- function(my_data, my_list) {
  for (each_item in my_list) {
  num_na <- sum(is.na(my_data[, each_item]))
  cat(each_item, ": ", num_na, '\n')
  }
}

# Call the missing_values function
missing_values(data, col_names)
```

The above shows the total number of missing values per variable.

```{r}
# Define a function to print the percentage of missing values for variables that have missing values
missing_values_pct <- function(my_data, my_list) {
  for (each_item in my_list) {
  num_na <- sum(is.na(my_data[, each_item]))
  if (num_na > 0) {
    pct_mv <- round((num_na / nrow(my_data)) * 100, 2)
    cat(each_item, ": ", pct_mv, '%', '\n')
  }
  }
}

# Call the missing_values_pct function
missing_values_pct(data, col_names)
```

The above shows the percentage of missing values for the variables that have missing values.\

Based on the code output above, it appears that the variables - ZVHI, inventory, median_list_price, median_sale_price, StL_ratio, prct_above_list, lat, and lng - all have missing values. Of the variables with missing values, ZVHI, inventory, median_list_price, median_sale_price, StL_ratio, and prct_above_list have percentages of missing values that exceed 5%.\

```{r total missing values}
# Make a copy of data
temp_data <- data.frame(data)

# Create a new column in temp_data for total missing values
temp_data$total_mv <- vector(mode="integer", length=nrow(temp_data))

# Define a function that adds 1 to a number
plus_one <- function(x) {
  x + 1
}

# Define a function that sums the number of total missing values per row
sum_mv <- function(my_data, col_names) {
  index_tracker <- 1
  for (each_row in 1:nrow(my_data)) {
  total_mv <- sum(is.na(my_data[each_row, col_names]))
  my_data$total_mv[index_tracker] <- total_mv
  index_tracker <- plus_one(index_tracker)
  }
  return(my_data)
}

# Call sum_mv function
temp_data <- sum_mv(temp_data, col_names)
str(temp_data)
```

```{r missing value plot}
# Plot date vs. missing values to see where we're missing the most values
ggplot(data = temp_data, aes(x=ts_month, y=total_mv)) + geom_bar(stat='identity')

# Another plot using data from 2017 onwards
ggplot(data = temp_data[temp_data$ts_month >= '2017-01-01', ], aes(x=ts_month, y=total_mv)) + geom_bar(stat='identity')
```

Based on the plots above, it looks like most of the missing values came from earlier years. We will adjust the time period of our analysis so that we focus on the year 2018 to the present-day.

```{r new data}
# Create a new dataframe with data from 2018-present day
clean_data <- data[data$ts_month >= '2018-01-01', ]

# View the new dataframe
str(clean_data)
```

```{r check new data for missing values}
# Call the missing_values_pct function to check the new dataframe
missing_values_pct(clean_data, col_names)
```

While median_sale_price, StL_ratio, and prct_above_list still have a high percentage of missing values, ZVHI, inventory, median_list_price, lat, and lng now have below 5% in missing values.

```{r save as csv}
# Write as csv file in Data folder
write.csv(clean_data, "../../Data/prelim_data_cleaning.csv")
```

### Appendix (Supplemental Code)

```{r unique values}
# Define a function to print the unique values for all variables
unique_values <- function(my_data, my_list) {
  for (each_item in my_list) {
  uv <- unique(my_data[, each_item])
  cat(each_item, ": ", uv, '\n')
  }
}

# Call the unique_values function - DO NOT RUN (too many output values)
#unique_values(data, col_names)
```

