---
title: "data_cleaning"
output: html_document
date: "2022-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Cleaning

The goal of this notebook is to add to my previous work on data cleaning.\

More specifically, I will handle missing values using various methods.

```{r env}
# Clear the environment
rm(list = ls())

# Set a random seed for code reproducibility
set.seed(25)
```

```{r libraries}
# Install required libraries
#install.packages("tidyr", repos = "http://cran.us.r-project.org")
#library(tidyr)
```

```{r data}
# Read the formatted csv file
data <- read.csv(file = "../../Data/prelim_data_cleaning.csv")

# View the structure of the data
str(data)
```

I will drop 3 columns since they don't contribute to the analysis: X.1, X, and ts_month.\

Note: I created a Datetime object called "ts_month" in a previous notebook, but I'm dropping it here since when I call the data file, it doesn't get read as a Datetime object.

```{r drop columns}
# Drop columns "X.1", "X", and "ts_month"
data <- subset(data, select = -c(X.1, X, ts_month))
str(data)
```

Let's review the percentage of missing values again.

```{r missing values}
# Create a list of column names
col_names <- colnames(data)

# Define a function to print the percentage of missing values for variables that have missing values
missing_values_pct <- function(my_data, my_list) {
  for (each_item in my_list) {
  num_na <- sum(is.na(my_data[, each_item]))
  if (num_na > 0) {
    pct_mv <- round((num_na / nrow(my_data)) * 100, 2)
    cat(each_item, ": ", pct_mv, '%', '\n')
  }
  }
}

# Call the missing_values_pct function
missing_values_pct(data, col_names)
```

There are 8 variables with missing values, 5 of which have less than 5% of missing values and 3 of which have greater than 40% of missing values.\

For our analysis, we are using the following as dependent variables: (1) prct_above_list, and (2) StL_ratio. All other variables will be used as predictors. It appears that our two response variables have the highest percentage of missing values.\

Before dropping any rows, we have 50512 observations in our data.\

**Method 1:** Drop all rows with missing values.

```{r method 1}
# Method 1: Drop all rows with missing values
data_1 <- data[complete.cases(data), ]
str(data_1)
```

```{r check na 1}
# Check missing values for data_1
col_names_1 <- colnames(data_1)
missing_values_pct(data_1, col_names_1)
```

Method 1 results in 31430 observations getting removed, leaving 19082 observations to analyze. If we were to use Method 1, we would keep 37.8% of the original data.\

**Method 2:** Drop all rows with missing values for the two response variables (i.e. prct_above_list and StL_ratio). In this case, dropping all null values for prct_above_list takes care of the null values for StL_ratio. The remaining missing values should then be imputed.

```{r method 2}
# Method 2: Drop all rows with missing values for the two response variables
data_2 <- data[!is.na(data$prct_above_list), ]
str(data_2)
```

```{r check na 2}
# Check missing values for data_2
col_names_2 <- colnames(data_2)
missing_values_pct(data_2, col_names_2)
```

Method 2 results in 28853 observations getting removed, leaving 21659 observations to analyze. If we were to use Method 2, we would keep 42.9% of the original data.\

**Method 3:** Change the timeframe to 2020-2022 and then drop all missing values. The year 2020 is selected because that is approximately when the two response variables appear in the data.

```{r method 3}
# Method 3: Change date to 2020-2022 and then drop all missing values
data_3 <- data[data$ts_month >= '2020-01-01', ]
str(data_3)
```

```{r check na 3}
# Check missing values for data_3
col_names_3 <- colnames(data_3)
missing_values_pct(data_3, col_names_3)
```

```{r method 3 drop nas}
# Drop all missing values
data_3 <- data_3[complete.cases(data_3), ]
str(data_3)
```

```{r check na 3 again}
# Check missing values for data_3 after dropping all missing values
col_names_3 <- colnames(data_3)
missing_values_pct(data_3, col_names_3)
```

Method 3 results in X observations getting removed, leaving 13425 observations to analyze. If we were to use Method 3, we would keep 26.6% of the original data.\

### Cleaned Data

Based on the results of the three methods outlined above, I will use Method 1 since it retains a decent percentage of the original data and doesn't require messy imputations for the data.

```{r save as csv}
# Write as csv file in Data folder
write.csv(data_1, "../../Data/data_cleaning.csv")
```